### 2020-05-07 chapter1.Rmd
"initial" commit

```{r libraries/definitions, include=FALSE}
library(stringi) # %s+%
library(lubridate) # floor_date
library(xts) # zoo
library(fpp) # auto.arima
library(tidyverse) # map_df, read_csv, group_by, summarize
library(tsbox) # https://cran.r-project.org/web/packages/tsbox/vignettes/tsbox.html
print_dec <- function(x, decimals=2) trimws(format(round(x, decimals), nsmall=decimals))

setwd("/media/data/Documents/Uni/spatiotemp/villalobos/project/")

# UCCS
domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"
# MEMS
# domain <- c(-7200,7200); pattern <- "MEM.*.csv"; name <- "MEM"; description <- "Membach"
```
```{r import dataset, include=FALSE}
seis <-
  list.files(path="csv/", pattern=pattern) %>%
  map_df(~read_csv("csv/" %s+% .))
```

# Data preparation
```{r dataset description, echo=FALSE}
long_description <- description %s+% " (" %s+% name %s+% " station)"
long_description
"Domain values: " %s+% domain[1] %s+% " to " %s+% domain[2] %s+% " nm"
"Data from " %s+% format.POSIXct(min(seis$utc_second), "%Y-%m-%d %H:%M:%S") %s+% " to " %s+% format.POSIXct(max(seis$utc_second), "%Y-%m-%d %H:%M:%S")
```

```{r cleanup, echo=FALSE}
out_domain <- (seis$min < domain[1] | seis$max > domain[2])
invalid_values <- (seis$min == -1 & seis$max == 1) | (seis$min == 0 & seis$max == 0) | out_domain
"Removing " %s+% sum(out_domain) %s+% " values out of domain, " %s+% (sum(invalid_values) - sum(out_domain)) %s+% " explicit NA"
seis <-
  seis %>%
  mutate(min=replace(min, invalid_values, NA), max=replace(max, invalid_values, NA)) %>%
  drop_na()
# %>% mutate(belgian_second=with_tz(utc_second, "Europe/Brussels"))
remove(out_domain, invalid_values) # free ~350 Mb from RAM
seis.present <- length(seis$utc_second)
seis.total <- max(as.numeric(seis$utc_second)) - min(as.numeric(seis$utc_second))
seis.missing <- seis.total - seis.present
seis.present %s+% " present values, " %s+% seis.missing %s+% " missing points (" %s+% print_dec(seis.present / seis.total * 100, 2) %s+% "% / " %s+% print_dec(seis.missing / seis.total * 100, 2) %s+% "%)"

```

```{r build_ts function}
build_ts <- function(dset, periods=NULL, start=NULL, end=NULL, force_ms=T) {
  dset.ts <- ts_zoo(dset)
  if (!is.null(start) | !is.null(end)) {
    dset.ts <- ts_span(dset.ts, start=start, end=end)
  }
  dset.ts <- ts_regular(dset.ts)
  if (is.null(periods)) {
    return(dset.ts)
  }
  if (is.vector(periods) & length(periods) == 1) {
    periods <- periods[1]
  }
  if (is.integer(periods)){
    if (force_ms) {
      periods <- c(periods)
    } else {
      frequency(dset.ts) <- periods
      return(dset.ts)
    }
  }
  return(msts(dset.ts, periods))
}
```

```{r group by hour}
seis.h <- 
  seis %>%
  mutate(utc_hour=floor_date(with_tz(utc_second, "Europe/Brussels"), "1 hour")) %>%
  group_by(utc_hour=as_datetime(date(utc_hour)) + hour(utc_hour) * 3600) %>% # this forces the hour to represent bruxelles local time in utc. tz-ed time is normally evaluated in utc by ts-related functions
  summarize(mean_movement_s=mean(max-min)) 
# we'll use ts_regular instead of completing data here
# %>% complete(utc_hour=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = as.difftime("1:0:0")))
```

```{r seasonplots}

seis.h.start <- with_tz(as_datetime('2018-11-01T00:00:00'), "Europe/Brussels")
seis.h.end <- with_tz(as_datetime('2020-03-01T00:00:00'), "Europe/Brussels")
seis.h.msts <- build_ts(seis.h, periods = c(24), start = seis.h.start, end = seis.h.end)
seasonplot(seis.h.msts, type="l", col=rainbow(7))
legend("topleft", legend=weekdays(seis.h.start + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
tsdisplay(seis.h.msts, lag.max = 168*2)
# tsdisplay(diff(seis.h.msts, 8), lag.max = 168*2)
tsdisplay(diff(seis.h.msts, 24), lag.max = 168*2)
tsdisplay(diff(seis.h.msts, 168), lag.max = 168*2)
tsdisplay(diff(diff(seis.h.msts, 24), 168), lag.max = 168*2)

seis.h.list <- as_datetime('2018-11-01T00:00:00') %m+% months(c(0:16))
for (i in 1:(length(seis.h.list)-1)) {
  seis.h.msts <- build_ts(seis.h, periods = c(24), start = seis.h.list[i], end = seis.h.list[i+1])
  seasonplot(seis.h.msts,
             type="l",
             xlab="Hour (" %s+% seis.h.list[i] %s+% " - " %s+% print_dec(length(seis.h.msts) / as.numeric(seis.h.list[i+1] - seis.h.list[i])) %s+% " points per day)",
             col=rainbow(7)
             )
  legend("topleft", legend=weekdays(seis.h.list[i] + (as.difftime(c(0:6), units="days"))), col=rainbow(7), lty=1, cex=0.8)
}

```

```{r test of ts, eval=FALSE, include=FALSE}
seis.h.xts <- build_ts(seis.h, start = seis.h.start, end = seis.h.end)
# https://stats.stackexchange.com/questions/120806/frequency-value-for-seconds-minutes-intervals-data-in-r
# https://robjhyndman.com/hyndsight/seasonal-periods/
seis.h.msts <- build_ts(seis.h, periods = c(24, 168), start = seis.h.start, end = seis.h.end)
taylor.fit <- tbats(seis.h.msts, num.cores = NULL)
taylor.fit
time(seis.h.msts)
plot(taylor.fit)
plot(forecast(taylor.fit))
ts_summary(seis.h.msts)
ts_plot(ts_trend(seis.h.xts))
ts_ggplot(seis.h.xts, title = 'Hourly seismic movement', subtitle = long_description %s+% " from " %s+% seis.h.start %s+% " to " %s+% seis.h.end, ylab='mean movement per second (nm)')
str(seis.h.xts)
plot(seis.h.xts)
seis.h.msts.dec <- decompose(seis.h.msts)
#seis.h.msts.dec$
plot(seis.h.msts.dec)
length(seis.h.xts)
frequency(seis.h.xts)
deltat(seis.h.xts)
```

```{r parallel fourier, include=FALSE}

combinations <- function(v1, v2=v1) {
  A <- matrix(nrow=0, ncol=2)
  for (i in v1) {
    for (j in v2) {
      A <- rbind(A, c(i, j))
    }
  }
  return(A)
}
auto.arima.fourier <- function(msts.data, periods, ...) {
  xreg.d <- fourier(msts.data, periods)
  print("Starting auto.arima with # of fourier factors " %s+% periods)
  fitma <- auto.arima(msts.data, ..., max.P=0, D=0, max.Q=0, seasonal=F, xreg=xreg.d)
  print("Finished auto.arima with # of fourier factors " %s+% periods)
  return(fitma)
}

library(parallel)
if (!is.null(getDefaultCluster())) {
  setDefaultCluster(makeCluster(6))
}

seis.h.start <- with_tz(as_datetime('2018-03-01T00:00:00'), "Europe/Brussels")
seis.h.end <- with_tz(as_datetime('2020-03-01T00:00:00'), "Europe/Brussels")
seis.h.msts <- build_ts(seis.h, periods = c(24, 168), start = seis.h.start, end = seis.h.end)
co <- combinations(7:12)
co <- co[order(rowSums(co)+co[,2], decreasing=T),]
clusterExport(getDefaultCluster(), c("seis.h.msts", "co", "auto.arima.fourier", "fourier", "auto.arima", "%s+%"))
fitma.list <- parLapply(getDefaultCluster(), 1:nrow(co), function(r) auto.arima.fourier(seis.h.msts, co[r,])) # TODO set arima parameters
stopCluster(getDefaultCluster())

aic_vals <- data.frame(Fourier24=integer(), Fourier168=integer(), AIC=integer())
for (fitma in fitma.list) {
  f24_l <- as.numeric(str_match(na.remove(str_extract(names(fitma$coef), "S\\d+-24")), "S(\\d+)-24")[,2])
  f24 <- max(f24_l)
  f168_l <- as.numeric(str_match(na.remove(str_extract(names(fitma$coef), "S\\d+-168")), "S(\\d+)-168")[,2])
  f168 <- max(f168_l)
  desc <- paste0("Seasonality component for ARMA(", fitma$arma[1], ",", fitma$arma[2], ")")
  subdesc <- paste0("Fourier terms [24h: ", length(f24_l), " - max ", f24, "] [168h: ", length(f168_l), " - max ", f168, "]")
  # plot(rowSums(fitma$xreg[,(c(1:(length(f24_l)*2)))][1:(168*2),]), type="l", col="red", ylim=c(-12,+24), main=desc)
  # lines(rowSums(fitma$xreg[,(c((length(f24_l)*2+1):(length(f168_l)*2+length(f24_l)*2)))][1:(168*2),]), type="l", col="blue")
  drift <- length(na.remove(str_extract(names(fitma$coef), "drift")))
  plot(rowSums(fitma$xreg[(1):(168*2),(drift+1):ncol(fitma$xreg)]), type="l", col="red", ylim=c(-12,+24), main=desc, sub=subdesc, xlab="Hours")
  aic_vals[nrow(aic_vals) + 1, ] <- c(length(f24_l), length(f168_l), fitma$aic)
}
p <- ggplot(aic_vals, aes(Fourier24, Fourier168)) + geom_tile(aes(fill = AIC), colour = "white") + scale_fill_gradient(low = "white", high = "steelblue")
plot(p)

min_aic <- min(aic_vals$AIC)
minvals <- aic_vals[which(aic_vals$AIC == min_aic),]
minvals

fitted <- fitma.list[order(aic_vals$AIC)][[2]]
tsdisplay(residuals(fitted))

```

```{r plot fourier}
str(fitted$xreg)

```

```{r fourier}
seis.h.start <- with_tz(as_datetime('2018-11-01T00:00:00'), "Europe/Brussels")
seis.h.end <- with_tz(as_datetime('2020-03-01T00:00:00'), "Europe/Brussels")
seis.h.msts <- build_ts(seis.h, periods = c(24, 168), start = seis.h.start, end = seis.h.end)
fitted <- auto.arima(seis.h.msts, D=0, max.P=0, max.Q=0, xreg=fourier(seis.h.msts, c(3, 4)))
plot(fitted)
plot(residuals(fitted))
tsdisplay(residuals(fitted))
res.xts <- ts_zoo(residuals(fitted))
res.msts <- msts(res.xts, c(24, 168))
res.aic_vals_temp <- NULL
res.aic_vals <- NULL
for (i in 1:5) {
  for (j in 1:5) {
    #for (k in 1:5) {
      res.xreg.d <- fourier(res.msts, c(i, j)  )  # daily  seasonality
      res.fitma <- auto.arima(res.msts, D=0, max.P=0, max.Q=0, xreg=res.xreg.d)
      res.aic_vals_temp <- cbind(i, j, res.fitma$aic)  #,k
      res.aic_vals <- rbind(res.aic_vals, res.aic_vals_temp)
      print("(" %s+% i %s+% ", " %s+% j %s+% ") = " %s+% res.fitma$aic)
    #}
  }
}
remove(res.aic_vals_temp, res.xreg.d, res.fitma)
colnames(res.aic_vals) <- c("Fourier24", "Fourier168", "AIC")
res.aic_vals <- data.frame(res.aic_vals)
res.min_aic <- min(res.aic_vals$AIC)
res.minvals <- res.aic_vals[which(res.aic_vals$AIC == res.min_aic),]
res.p <- ggplot(res.aic_vals, aes(Fourier24, Fourier168)) + geom_tile(aes(fill = AIC), colour = "white") + scale_fill_gradient(low = "white", high = "steelblue")
plot(res.p)
res.minvals
res.fitted <- auto.arima(res.msts, D=0, max.P=0, max.Q=0, xreg=fourier(res.msts, c(res.minvals$Fourier24, res.minvals$Fourier168)))
plot(res.fitted)
tsdisplay(residuals(fitted))
tsdisplay(residuals(res.fitted))

```

```{r periodgram}
library(spec)
#pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2018-11-01T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)
pgram <- spec.pgram(build_ts(seis.h, start=as_datetime('2019-05-03T00:00:00'), end=as_datetime('2020-03-01T00:00:00')), main="periodgram", detrend=F, demean=T, na.action=na.aggregate)

topfreq <- pgram$freq[order(pgram$spec, decreasing = T)][1:100]
topseas <- 1/topfreq
topseas.h <- topseas/3600

top95pc <- pgram$spec[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*1)]
top95freq <- pgram$freq[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*1)][order(top95pc, decreasing = T)]
top95pc <- pgram$spec[pgram$spec > (mean(pgram$spec) + sd(pgram$spec)*1)][order(top95pc, decreasing = T)]
top95seas.s <- 1/top95freq
top95seas.h <- top95seas.s/3600
top95pc/sd(pgram$spec)
top95seas.h

ylims <- 10^par("usr")[3:4]
axis(side=3, at=top95freq, labels=as.integer(top95seas.h), padj = 1, las=2, cex.axis=0.8)
abline(h=mean(pgram$spec) + sd(pgram$spec)*c(.5,1:3), lty="dotted", col=c("orange", "darkgreen", "blue", "red"))
for (i in 1:length(top95freq)) {
  color <- ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*3), "red", ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*2), "blue", ifelse(top95pc[i] > (mean(pgram$spec) + sd(pgram$spec)*1), "darkgreen", "orange")))
  lines(x=c(top95freq[i], top95freq[i]), y=c(top95pc[i],ylims[2]), lty="dotted", col=color)
  points(x=top95freq[i], y=top95pc[i], col=color)
}
axis(side=4, at=mean(pgram$spec) + sd(pgram$spec) * c(.5, 1:3), labels=c(".5", "1", "2", "3"), cex.axis=0.8)
mtext(side=4, "spectrum sd ratio", line=1)
```