### 2020-05-22 project.uccs.Rmd
more analysis and better Rmd output

---
title: "Project Uccle"
author: "Marco Ciotola"
output:
  pdf_document:
    keep_tex: yes
    fig_width: 6.5
    fig_height: 4
#  html_document:
#    mathjax: local
#    self_contained: no
#    fig_width: 6.5
#    fig_height: 4
---

```{r tentatives, eval=FALSE, child='chapter1.Rmd', include=FALSE}
```

```{r set uccs, include=FALSE}
# UCCS
domain <- c(-350000,350000); pattern <- "UCCS.*\\.csv"; name <- "UCCS"; description <- "Uccle, Bruxelles"; proj_file <- 'proj.uccs.RData'
```

```{r project, child='project.Rmd'}
```